### 2020-12-14 ch/1-introduction.tex
Some modest progress

This problem becomes especially severe when the training data is biased in some way,
which, by default, makes the \ac{ML} system copy the bias, or even exacerbate it.
The topic of this thesis is how to deal with biased data;
what types of bias exist and how to correct it if possible.

\section{Motivation and aims}

>>>>

This problem becomes especially severe when the training data is biased in some way,
which, by default, makes the \ac{ML} system copy the bias, or even exacerbate it.
When applied in the deployment setting, the system will then not behave in the desired way.
The topic of this thesis is how to deal with biased data;
what types of bias exist and how to correct it if possible.

\section{Motivation and aims}
Many datasets with person-related features display biases when examined by common fairness criteria.
This can range from relatively harmless biases,
like men being, on average, older than women in the CelebA dataset \citep{liu2015faceattributes},
to more serious ones,
like black men being several times less likely to receive bail than white men in the COMPAS dataset \citep{angwin2016machine}.
In the absence of truly `fair' datasets, our methods have to be able to deal with these biases.

Now, one possible objection here is:
if those datasets are so bad,
then maybe we just should not train any \ac{ML} model on these and should not use them to make automated decisions.
While this question is mostly beyond the scope of this document, let me offer some thoughts on this:
It's true that even after the application of de-biasing techniques,
the resulting models still should not be fully trusted,
but they can still ease the burden of checking everything manually;
similar to an email spam filter which is not perfect, but still very useful.
One could imagine a hybrid approach where an automated system makes preliminary decisions,
but random samples are reviewed by humans and decisions can always be challenged.
Ideally, the model itself would tell us about decisions it is uncertain about.

Moreover, two of the three methods presented in this thesis require access to (unlabeled) \emph{unbiased} data,
which is used during the training process.
So, the criticism that we are only learning from biased data does not apply here.
This should allow us to be more confident in the predictions of those methods.

Throughout this thesis, the aim is to learn a model from biased training data,
which gives fair predictions on an unbiased training set.
It is thus not enough to simply optimize the cross-entropy on the training set;
we have to change the optimization target to achieve our stated goal.
While it is possible to extend notions of fairness beyond classification tasks,
the vast majority of work in this area concerns classification only and that will be the case here as well.

\section{Claims and contributions}%
\label{sec:claims-contributions}
This thesis is based on 3 publications.
\begin{enumerate}
  \item x
\end{enumerate}
This thesis will discuss two kinds of dataset bias: \emph{label bias} and \emph{sampling bias}.

In the setup for \emph{label bias},
we consider a classification problem in which the class labels $y$ needs to be predicted from input features $\vx$.
Additionally, there is a \emph{sensitive attribute} $s$ associated with each input.
This sensitive attribute usually encodes membership in a demographic group, such as gender,
but more generally is a feature that should not be used to make predictions for legal or ethical reasons.
In our examples, $s$ and $y$ are binary variables, but this does not have to be the case.
The label bias now manifests itself in a specific way:
depending on $s$, labels are either flipped from $y=0$ to $y=1$ or vice versa,
\ie, there is an error in the labels which is correlated with the sensitive attribute.

---

I'm not sure how to mention my papers here. First the papers and then the description?
